How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-01 tse-paper/content.tex
address all of the todos addressed to Caro

The results of our empirical comparison study on the chunk retrieval techniques
show that the structural representation of the targeted chunk greatly influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example-1,lst:chunk-e:xample-2} show two exemplary log chunks
from Android/Gradle build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed. When PBE is trained on these two examples it produces
a regular expression similar to \Cref{lst:prose-program-simplified}. The retrieval
starts after a colon followed by a line separator and before a capital letter.
It ends before two line separators, consistent with the two example chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2}.
This shows, that the characters structuring the text around the chunk are
cruicial for regular expressions to identify the targeted log chunk.
In fact, if the third
example from \Cref{lst:chunk-example-3} with a different structural
representation is added, PBE is not able to syntesize a program and returns
``\texttt{no program found}''.
We express this difference in structural representation through dividing log
chunks into structural categories. The chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2} are in the same structural
category, while the chunk from \Cref{lst:chunk-example-3} is in a different one.

@@

% TODO (Caro): Integrate Kitchenham's guidelines
From our own experience, we know that the field of analyzing build
logs is relatively young, with perhaps the work on TravisTorrent the
most cited and central one. We also know there have been scattered
research efforts on build log analysis prior to it, mostly where the
log analysis was a side-part of the article. To the best of our
knowledge, there has not been an attempt to systematically overview
and survey them, despite their forseeable future and current importance.

@@

With these 259 remaining articles, we followed a backward snowball
sampling for related work. This added three papers (five before
duplicate elimination) to our literature set, coming to a total of 261
articles. % TODO Caro: update

@@

\subsection{Article Classification}
% TODO Caro: please write a draft for this :-)
Having trimmed down the set of articles, we investigated them in depth.
kept all articles that use logs from software builds as input to some kind of process.
we wanted to characterize how researchers use buildlogs as input to any processes.
Interesed in 
\begin{itemize}
  \item what kind of build logs they are using and the source of these logs.
  \item what information they are retrieving from the logs, 
  its kind (chunk/classification) and 
  what they use the retrieved information for
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their tools
\end{itemize}
To objecifiy our data extraction we created a template of the targeted factors
and performed a pilot on five articles. Two authors extracted the data from the
five pilot articles and we held a consensus meeting to unify our understanding
of the factors in the data extraction templates. The remaining articles were
divided over these two authors. We inspected the articles as closesly as necessary,
starting from occurences of ``log'' or ``build'' within the text.
The full data extraction template and our results of this step can be found in
our replication package \todo{cite}.

@@

@@

\subsubsection{Configuration}
To configure chunk retrieval though text similarity we chose to use
the same concept of examples as for PBE\@
% TODO Caro: More details on this. How do you 'apply' VSM? On what?
and apply the Vector Space Model~\cite{schutze2008introduction}. The
lines of the output strings of the training examples define our search
query. The algorithm splits the search query into single lines and
identifies tokens, in our case words. Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words. Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@. To better
compare KWS with PBE and CTS, we also configure it through examples.
We associate each example with keywords which appear in the targeted
chunk or close to it. The configuring keywords for KWS are the ones
that appear most often in the keywords of all training examples.
% TODO Caro Please explain above sentence more precisely

@@

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example, we refer to them in short as \emph{examples}. The
\emph{input} is the text of the build log file. The \emph{output} is
an array of substrings of the log file text, representing the
substrings that should be retrieved by the synthesized program when
given the corresponding input file. One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted. The
PROSE program synthesis then tries to construct a program consistent
with all training examples.
% A program is consistent with an example if it returns the defined
% output when executed on the defined
% input~\cite{mitchell1982generalization}.
PBE reports an error back to the user if it could not synthesize a
consistent program.
% The program synthesis builds on the FlashExtract DSL, which in turn
% uses the FlashMeta algorithm. Both are described in
% Section~\ref{sec:rw-prose}.

% TODO Caro: The following is very obvious and could apply to any
% technique. 
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program. It then returns the substring
of the build log matched by the program or an empty string if the
program found no match.

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}

In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log. RLR essentially mimicks the
situation of guessing purely blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return. For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples. 
% TODO Caro: some more details here?

@@

@@

\noindent
\textbf{Training and Test Set}
We use \emph{LogChunks} as the data set for our study. For each of the
80 repositories in it, \emph{LogChunks} contains 10 build logs,
manually labeled with the substring or \emph{chunk} describing the
reason the build failed. Additionally, it contains (1) keywords useful
to
% TODO Caro uniquely?
identify the substring in the log  and (2) which structural category
the substring belongs to.
% TODO (Caro): Add details how key words were extracted. Add example?

% TODO (Caro): Add more details. How was the split in numbers?
For each repository in \emph{LogChunks}, we split the examples
chronologically into training and test set. Therefore, we train on
examples from past build logs and test on more recent build logs.

\noindent
\textbf{RQ 2.1: Size of Training and Test Set}
To analyze how many examples the chunk retrieval techniques need to
perform best, we evaluate the techniques with different training set
sizes. We train each technique with one to five examples from each of
the repositories within \emph{LogChunks}. The size of the test set is
one.

@@

\vspace{0.2cm}
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
	\item $|\mbox{True positives}| = \mathit{DesiredLines} \cap
	\mathit{RetrievedLines}$ \vspace{0.2cm}\\
        True positives are lines that appear both in the output of a
        tool as well as what \textit{LogChunks} contains for each
        example. 
        % TODO (Caro) What about when a line is replicated twice?
        % I.e., Are line numbers part of this?
	\item $\mbox{Precision} = \dfrac{|\mathit{True\
	Positives}|}{|\mathit{RetrievedLines}|}$ \vspace{0.21cm} \\
        Precision of a chunk retrieval describes which proportion of
        the retrieved lines were actually desired. 
	\item $\mbox{Recall} =
	\dfrac{|\mathit{TruePositives}|}{|\mathit{DesiredLines}|}$
	\vspace{0.2cm} \\
        Recall of a chunk retrieval describes which proportion of the
        desired lines were retrieved.
	\item $\mbox{F$_{1}$-score} = 2 \cdot \dfrac{\mathit{Precision}
	\cdot \mathit{Recall}}{\mathit{Precision} + \mathit{Recall}}$
	\vspace{0.2cm}\\
        In addition, we calculate the F$_{1}$-score, the harmonic mean
        of precision and recall. We prefer F$_{1}$ to other aggregate
        measures such as accuracy because for a
        ``needle-in-the-haystack'' scenario, we do not want to bloat
        our results by correctly not finding lots of irrelevant log
        lines.
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}$  \vspace{0.2cm} \\
	We define a successful retrieval as one where all desired lines were
	extracted, therefore when recall is one.
\end{itemize}

@@

In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log. Due to the fact no one would
seriously consider RLR in real-world scenarios, we are omitting its
detailed graphs here. However, its results follow intuitive
expectations:
% TODO Caro: shortly descibre RLR graphs ommitted

@@

% TODO Rather than giving them once (``friss und stirb''), better to incorporate these RQs into the text where they are being answered

@@

% TODO Caro: I do not understand the following sentence.
Extracting the average number of lines present in the training
examples gives the best retrieval results for CTS, according to the
F$_{1}$-score.

\noindent
\textbf{Retrieval Output Usage}
% TODO Caro: reforumlate this please
CTS has good precision and recall on average, though the precision or
recall of a chunk retrieval run is very hard to predict from the given
result. Therefore, retrieval output from CTS is suited to be read by a
human. This could include semi-automated procedures such as sending
developers an email with the extracted build failure reason.

@@

\noindent
\textbf{Few Samples with Many Structural Categories}
Our comparison study shows fewer measurements with many structural
categories than with one category.
% TODO Caro: Please explain the following further. Why does a
% realistic data set only have one or few structural categories?
This stems from the fact that we use a realistic data set which in
many cases has only one or few structural categories.

@@

There are various future research opportunities based on our work:
\begin{itemize}
  \item \textbf{Further Analysis of \emph{LogChunks}} We created the
  \emph{LogChunks} data set \cite{} specifically for the comparative
  study in this paper, though it can be the basis for various further
  analyses of build log data. The keywords, for example, can be
  investigated to answer which keywords are used to search for the
  reason the build failed within build logs.
  \item \textbf{Cross-Repository Build Log Analysis} We trained and
  tested each chunk retrieval technique on examples from the same
  repository. We propose to analyze how techniques could be trained
  across repositories, building the cornerstones for build
  environment-agnostic analysis tools.
  \item \textbf{Comparison with more Chunk Retrieval Techniques} This
  paper investigates the three chunk retrieval techniques PBE, CTS and
  KWS\@. Our study design can be reused to evaluate other build log
  analysis techniques, such as the diff and information retrieval
  approach by Amar et al.~\cite{amar2019mining}.
  \item \textbf{Refinement of Retrieval Quality for each Technique} We
  investigated basic configurations of existing techniques applied to
  chunk retrieval from build logs. In a next step, each of these
  techniques could be refined to better approach the domain of build
  logs. The \emph{LogChunks} data set and our study results act as a
  baseline to benchmark such technique improvements. We propose the
  following improvements:
    \begin{itemize}
      \item \textbf{Custom Ranking and Tokens for PBE} The program
      synthesis through PROSE ranks possible programs according to
      what the user most likely intended. One could adapt the ranking
      rules provided by the FlashExtract DSL to fit common build log
      chunk retrieval tasks. FlashExtract includes special tokens when
      enumerating possible regular expressions. One could extend these
      with tokens found in build logs, such as ``-'',``='',``ERROR''
      or ``[OK''.
      \item \textbf{Meta-Parameter Optimization for CTS} Information
      retrieval techniques have various meta-parameters which can be
      optimized for the specific use
      case~\cite{panichella2016parameterizing}. We propose to further
      investigate improvements in preprocessing of the log text, in
      tokenization of the log lines into terms and in stop words
      lists.
    \end{itemize}
  \item \textbf{Usability Analysis of Chunk Retrieval Output} Our
  analysis of the output produced by the chunk retrieval focuses on
  precision and recall. We propose to investigate how useful these
  outputs are to developers through controlled experiments.
\end{itemize}

>>>>

The results of our empirical comparison study on the chunk retrieval techniques
show that the structural representation of the targeted chunk greatly influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example-1,lst:chunk-example-2} show two exemplary log chunks
from Android/Gradle build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed. When PBE is trained on these two examples it produces
a regular expression similar to \Cref{lst:prose-program-simplified}. The retrieval
starts after a colon followed by a line separator and before a capital letter.
It ends before two line separators, consistent with the two example chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2}.
This shows, that the characters structuring the text around the chunk are
cruicial for regular expressions to identify the targeted log chunk.
In fact, if the third
example from \Cref{lst:chunk-example-3} with a different structural
representation is added, PBE is not able to syntesize a program and returns
``\texttt{no program found}''.
We express this difference in structural representation through dividing log
chunks into structural categories. The chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2} are in the same structural
category, while the chunk from \Cref{lst:chunk-example-3} is in a different one.

@@

% TODO (Caro): Integrate Petersen's guidelines
From our own experience, we know that the field of analyzing build
logs is relatively young, with perhaps the work on TravisTorrent the
most cited and central one. We also know there have been scattered
research efforts on build log analysis prior to it, mostly where the
log analysis was a side-part of the article. To the best of our
knowledge, there has not been an attempt to systematically overview
and survey them, despite their forseeable future and current importance.

@@

With these 259 remaining articles, we followed a backward snowball
sampling for related work. This added three papers (five before
duplicate elimination) to our literature set, coming to a total of 261
articles. % TODO: update numbers

@@

\subsection{Article Classification}
Having trimmed down the set of articles, we investigated them in depth.
kept all articles that use logs from software builds as input to some kind of process.
we wanted to characterize how researchers use buildlogs as input to any processes.
Interesed in 
\begin{itemize}
  \item what kind of build logs they are using and the source of these logs.
  \item what information they are retrieving from the logs, 
  its kind (chunk/classification) and 
  what they use the retrieved information for
  \item which technique they are using to process the build logs,
  in how much detail it is described and whether they published their tools
\end{itemize}
To objecifiy our data extraction we created a template of the targeted factors
and performed a pilot on five articles. Two authors extracted the data from the
five pilot articles and we held a consensus meeting to unify our understanding
of the factors in the data extraction templates. The remaining articles were
divided over these two authors. We inspected the articles as closesly as necessary,
starting from occurences of ``log'' or ``build'' within the text.
The full data extraction template and our results of this step can be found in
our replication package \todo{cite}.

@@

% TODO later: we should make sure to at least mention PROSE
% _somewhere_, it's crucial for why PBE is how it is
\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example, we refer to them in short as \emph{examples}. The
\emph{input} is the text of the build log file. The \emph{output} is
an array of substrings of the log file text, representing the
substrings that should be retrieved by the synthesized program when
given the corresponding input file. One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted. The
PROSE program synthesis then tries to construct a program consistent
with all training examples.
% A program is consistent with an example if it returns the defined
% output when executed on the defined
% input~\cite{mitchell1982generalization}.
PBE reports an error back to the user if it could not synthesize a
consistent program.
% The program synthesis builds on the FlashExtract DSL, which in turn
% uses the FlashMeta algorithm. Both are described in
% Section~\ref{sec:rw-prose}.

% TODO Moritz: you said: The following is very obvious and could apply to any
% technique. 
% kinda. It does give a descriptive error message on no match (and
% that is not possible in other techniques)
% otherwise yes, nothing interesting happens in the application of pbe
% though for consistency we might need this section? or merge config/application?
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program. It then returns the substring
of the build log matched by the program or an error message if the
program found no match.

@@

\subsubsection{Configuration}
To configure chunk retrieval though text similarity we chose to use
the same concept of examples as for PBE
% TODO Moritz: your question was: More details on this. How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM. word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation? 
and apply the Vector Space Model~\cite{schutze2008introduction} as follows. The
lines of the output strings of the training examples define our search
query. The algorithm splits the search query into single lines and
identifies tokens, in our case words. Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words. Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@. To better
compare KWS with PBE and CTS, we also configure it through examples.
We associate each example with keywords which appear in the targeted
chunk or close to it. 
KWS then searches for those keywords which are tied to the greatest
number of examples in the training set.
If several keywords are associated with the same number of training
examples, and no other keywords are associated with more training
examples, KWS searches for all of these keywords.

@@

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}
In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log. RLR essentially mimicks the
situation of guessing purely blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return. For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples. 
% TODO Moritz: you said: some more details here?
% I would not know of any details to add. It's really all we are doing

@@

% TODO: we could add more techniques from the lit. survey here

@@

\noindent
\textbf{Training and Test Set}
We use \emph{LogChunks} as the data set for our study. For each of the
80 repositories in it, \emph{LogChunks} contains 10 build logs,
manually labeled with the \emph{chunk} describing the
reason the build failed. Additionally, it contains (1) keywords useful
to identify the substring in the log  and (2) which structural category
the substring belongs to.
% TODO Moritz: You said: Add details how key words were extracted. Add example?
% how the keywords were put in logchunks does not fit here, that is
% maybe answered 2 paragraphs above already?
% how KWS uses the keywords is already described in the section
% introducing the technique => I would scrap everything but the first
% sentence directly above here :)

% TODO Moritz: You said: Add more details. How was the split in numbers?
% this is described in the following paragraph, should we merge them
% (bc there are also no sub-RQs now) or is it close enough? :)
For each repository in \emph{LogChunks}, we split the examples
chronologically into training and test set. Therefore, we train on
examples from past build logs and test on more recent build logs.

\noindent % TODO: adapt if we change the research questions
\textbf{RQ 2.1: Size of Training and Test Set}
To analyze how many examples the chunk retrieval techniques need to
perform best, we evaluate the techniques with different training set
sizes. We train each technique with one to five examples from each of
the repositories within \emph{LogChunks}. The size of the test set is
one.

@@

\vspace{0.2cm}
\begin{itemize}[leftmargin=0.4cm] \itemsep1em
	\item $|\mbox{True positives}| = \mathit{DesiredLines} \cap
	\mathit{RetrievedLines}$ \vspace{0.2cm}\\
        True positives are lines that appear both in the output of a
        tool as well as what \textit{LogChunks} contains for each
        example. 
        % What about when a line is replicated twice?
        % I.e., Are line numbers part of this? => no.
        % not checked, but if lines are identical they also contain
        % the same information so if someone asks we can defend I think
	\item $\mbox{Precision} = \dfrac{|\mathit{True\
	Positives}|}{|\mathit{RetrievedLines}|}$ \vspace{0.21cm} \\
        Precision of a chunk retrieval describes which proportion of
        the retrieved lines were actually desired. 
	\item $\mbox{Recall} =
	\dfrac{|\mathit{TruePositives}|}{|\mathit{DesiredLines}|}$
	\vspace{0.2cm} \\
        Recall of a chunk retrieval describes which proportion of the
        desired lines were retrieved.
	\item $\mbox{F$_{1}$-score} = 2 \cdot \dfrac{\mathit{Precision}
	\cdot \mathit{Recall}}{\mathit{Precision} + \mathit{Recall}}$
	\vspace{0.2cm}\\
        In addition, we calculate the F$_{1}$-score, the harmonic mean
        of precision and recall. We prefer F$_{1}$ to other aggregate
        measures such as accuracy because for a
        ``needle-in-the-haystack'' scenario, we do not want to bloat
        our results by correctly not finding lots of irrelevant log
        lines.
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}$  \vspace{0.2cm} \\
	We define a successful retrieval as one where all desired lines were
	extracted, therefore when recall is one.
\end{itemize}

@@

In our evaluation, we compare against a baseline of randomly
picking lines from the build log. Its results follow intuitive
expectations:
The precision ranges between 8\% and 5\%, the recall between 6\% and
8\%. The size of the training set has no visible impact on either.
An greater number of structural categories within the training and
test sets decreased the precision of RLR from 7\% to 0\%, the recall
increases from 6\%, for one structural category present, to 11\%, for
three sturctural categories present and also drops to 0\% for more
structural categories.
Graphs of the results of the chunk retrieval with RLR from our study
are included in our replication package~\cite{brandt2020chunk-replication}.

@@

% TODO update RQs
% TODO incorporate these RQs into the text where they are being answered

@@

\noindent
\textbf{Retrieval Output Usage}
CTS has good precision and recall on average, though with a high
variation. This means that the quality of an output by CTS is hard to
determine, which makes it unsuited for automatic  processing and requires a human to further inspect and interpret the output. This could include semi-automated procedures such as sending
developers an email with the extracted build failure reason.

@@

\noindent
\textbf{Few Samples with Many Structural Categories}
Our comparison study shows fewer measurements with many structural
categories than with one category.
% TODO Caro: Please explain the following further. Why does a
% realistic data set only have one or few structural categories?
This stems from the fact that we investigated the chunk retrieval
techniques on a realistic data set, which showed to often have few
structural categories within one project.

@@

There are various future research opportunities based on our work:
\begin{itemize}
  \item \textbf{Further Analysis of \emph{LogChunks}} We created the
  \emph{LogChunks} data set \cite{brandt2020logchunks} specifically for the comparative
  study in this paper, though it can be the basis for various further
  analyses of build log data. The keywords, for example, can be
  investigated to answer which keywords are used to search for the
  reason the build failed within build logs.
  \item \textbf{Cross-Repository Build Log Analysis} We trained and
  tested each chunk retrieval technique on examples from the same
  repository. We propose to analyze how techniques could be trained
  across repositories, building the cornerstones for build
  environment-agnostic analysis tools.
  \item \textbf{Comparison with more Chunk Retrieval Techniques} This
  paper investigates the three chunk retrieval techniques PBE, CTS and
  KWS\@. Our study design can be reused to evaluate other build log
  analysis techniques, such as the diff and information retrieval
  approach by Amar et al.~\cite{amar2019mining}.
  \item \textbf{Refinement of Retrieval Quality for each Technique} We
  investigated basic configurations of existing techniques applied to
  chunk retrieval from build logs. In a next step, each of these
  techniques could be refined to better approach the domain of build
  logs. The \emph{LogChunks} data set and our study results act as a
  baseline to benchmark such technique improvements. We propose the
  following improvements:
    \begin{itemize}
      \item \textbf{Custom Ranking and Tokens for PBE} The program
      synthesis through PROSE ranks possible programs according to
      what the user most likely intended. One could adapt the ranking
      rules provided by the FlashExtract DSL to fit common build log
      chunk retrieval tasks. FlashExtract includes special tokens when
      enumerating possible regular expressions. One could extend these
      with tokens found in build logs, such as ``-'',``='',``ERROR''
      or ``[OK''.
      \item \textbf{Meta-Parameter Optimization for CTS} Information
      retrieval techniques have various meta-parameters which can be
      optimized for the specific use
      case~\cite{panichella2016parameterizing}. We propose to further
      investigate improvements in preprocessing of the log text, in
      tokenization of the log lines into terms and in stop words
      lists.
    \end{itemize}
  \item \textbf{Usability Analysis of Chunk Retrieval Output} Our
  analysis of the output produced by the chunk retrieval focuses on
  precision and recall. We propose to investigate how useful these
  outputs are to developers through controlled experiments.
\end{itemize}
